```{r}
chess_df <- read.csv("analysis.csv")
```
##Question 1 a
Select useful data.
```{r}
df <- subset(chess_df,White.Player=='Anand, Viswanathan'|White.Player=='Carlsen, Magnus')
df <- df[,c('White.Player','White.ACPL')]
df
```
Summary of data and standard deviation.
```{r}
summary(df)
sd(df$White.ACPL)
```
Plot box plot.
```{r}
library(ggplot2)
ggplot(df) + geom_boxplot(aes(White.Player, White.ACPL, fill = White.Player)) + geom_jitter(aes(White.Player, White.ACPL, shape = df$White.Player))
```
Calculate the mean, median and standard deviation of ACPL for each player.
```{r}
tapply(df$White.ACPL, df$White.Player, mean)
```
```{r}
tapply(df$White.ACPL, df$White.Player, median)
```
```{r}
tapply(df$White.ACPL, df$White.Player, sd)
```
Perform t-test
```{r}
t.test(White.ACPL ~ White.Player, data=df, var.equal = TRUE)
```
```{r}
compare_2_gibbs <- function(y, ind, maxiter = 10000)
{
  mu0 = 12.5
  tau0 = 1/(10^2)
  del0 = 0
  gamma0 = 1/(10^2)
  a0 = 1
  b0 = 10^2
  
  y1 <- y[ind == 1]
  y2 <- y[ind == 2]

  n1 <- length(y1) 
  n2 <- length(y2)

  # starting values
  mu <- (mean(y1) + mean(y2)) / 2
  del <- (mean(y1) - mean(y2)) / 2
  

  mat_store <- matrix(0, nrow = maxiter, ncol = 3)
  
  an <- a0 + (n1 + n2)/2
  
  for(s in 1 : maxiter) 
  {
  
    # sample tau
    bn <- b0 + 0.5 * (sum((y1 - mu - del) ^ 2) + sum((y2 - mu + del) ^ 2))
    tau <- rgamma(1, an, bn)
  
    # sample mu
    taun <-  tau0 + tau * (n1 + n2)
    mun <- (tau0 * mu0 + tau * (sum(y1 - del) + sum(y2 + del))) / taun
    mu <- rnorm(1, mun, sqrt(1/taun))
  
    # sample del
    gamman <-  gamma0 + tau*(n1 + n2)
    deln <- ( del0 * gamma0 + tau * (sum(y1 - mu) - sum(y2 - mu))) / gamman
    del<-rnorm(1, deln, sqrt(1/gamman))
  
    # store values
    mat_store[s, ] <- c(mu, del, tau)
  }
  colnames(mat_store) <- c("mu", "del", "tau")
  return(mat_store)
}

library(MCMCpack)
df$White.Player[df$White.Player == "Anand, Viswanathan"] = 1
df$White.Player[df$White.Player == "Carlsen, Magnus"] = 2
fit <- compare_2_gibbs(df$White.ACPL, as.factor(df$White.Player))
plot(as.mcmc(fit))
```
Diagnosis of the sampler.
```{r}
raftery.diag(as.mcmc(fit))
```
Get summary, know the parameters of posterior distribution. 
```{r}
apply(fit, 2, mean)
```
```{r}
apply(fit, 2, sd)
```
Generate posterior for each player and plot the difference between mean of ACPL.
```{r}
player1 <- rnorm(10000, fit[,1]+fit[,2], sd=1/sqrt(fit[, 3])) 
player2 <- rnorm(10000, fit[,1]-fit[,2], sd=1/sqrt(fit[, 3])) 
ggplot(data.frame(diff = player1 - player2)) + stat_bin(aes(diff))
```
```{r}
ggplot(data.frame(player1, player2)) + geom_point(aes(player1, player2), alpha = 0.3) + geom_abline(slope = 1, intercept = 0)
```
Probability of one player is better than the other.
```{r}
mean(player1 > player2)
```
##Question 1 b
Select useful data
```{r}
white_df <- chess_df[,c('White.Player','White.ACPL','White.Player_ID')]
white_df
```
Summary of data and standard deviation.
```{r}
summary(white_df)
sd(white_df$White.ACPL)
```

Visualize the data
```{r}
white_df$White.Player <- factor(white_df$White.Player)
nlevels(white_df$White.Player)
ggplot(white_df) + geom_boxplot(aes(x = reorder(White.Player, White.ACPL, median), White.ACPL, 
                               fill = reorder(White.Player, White.ACPL, median)), show.legend=FALSE) +
                              theme(axis.text.x = element_text(angle = 90))
```
```{r}
ggplot(white_df, aes(x = reorder(White.Player, White.Player, length))) + stat_count() + theme(axis.text.x = element_text(angle = 90))
```
```{r}
ggplot(white_df, aes(White.ACPL)) + stat_bin()
```
```{r}
ggplot(data.frame(size = tapply(white_df$White.ACPL, white_df$White.Player, length), 
                  mean_score = tapply(white_df$White.ACPL, white_df$White.Player, mean)), 
       aes(size, mean_score)) + geom_point()
```
```{r}
compare_m_gibbs <- function(y, ind, maxiter = 10000)
{
  
  mu0 = 20
  tau0 = 1/100
  a0 = 1
  b0 = 625
  alpha0 = 1
  beta0 = 625
  
  # starting values
  m <- nlevels(ind)
  ybar <- theta <- tapply(y, ind, mean)
  tau_w <- mean(1 / tapply(y, ind, var)) 
  mu <- mean(theta)
  tau_b <-var(theta)
  n_m <- tapply(y, ind, length)
  alpha_n <- alpha0 + sum(n_m)/2
  
  theta_mat <- matrix(0, nrow=maxiter, ncol=m)
  mat_store <- matrix(0, nrow=maxiter, ncol=3)
  
  for(s in 1:maxiter) 
  {
    
    # sample thetas
    for(j in 1:m) 
    {
      taun <- n_m[j] * tau_w + tau_b
      thetan <- (ybar[j] * n_m[j] * tau_w + mu * tau_b) / taun
      theta[j]<-rnorm(1, thetan, 1/sqrt(taun))
    }
    
    # sample tau_w
    ss <- 0
    for(j in 1:m){
      ss <- ss + sum((y[ind == j] - theta[j])^2)
    }
    beta_n <- beta0 + ss/2
    tau_w <- rgamma(1, alpha_n, beta_n)
    
    # sample mu
    tau_m <- m * tau_b + tau0
    mu_m <- (mean(theta) * m * tau_b + mu0 * tau0) / tau_m
    mu <- rnorm(1, mu_m, 1/ sqrt(tau_m)) 
    
    # sample tau_b
    a_m <- a0 + m/2
    b_m <- b0 + sum((theta - mu)^2) / 2
    tau_b <- rgamma(1, a_m, b_m)
     
    # store values
    theta_mat[s,] <- theta
    mat_store[s, ] <- c(mu, tau_w, tau_b)
  }
  colnames(mat_store) <- c("mu", "tau_w", "tau_b")
  colnames(theta_mat) <- levels(ind)
  return(list(params = mat_store, theta = theta_mat))
}

fit2 <- compare_m_gibbs(white_df$White.ACPL, as.factor(white_df$White.Player_ID))
plot(as.mcmc(fit2$params))
```
Diagnosis of the sampler.
```{r}
raftery.diag(as.mcmc(fit2$params))
```

```{r}
apply(fit2$params, 2, mean)
apply(fit2$params, 2, sd)
mean(1/sqrt(fit2$params[, 3]))
sd(1/sqrt(fit2$params[, 3]))
```
```{r}
# reformat samples
samples_df <- data.frame(samples = as.numeric(fit2$theta), 
                       player = rep(1:ncol(fit2$theta), each = nrow(fit2$theta))) 

samples_mean <- apply(samples_df, 2, mean)
sort(samples_mean, decreasing = TRUE) 
```
```{r}
ggplot(samples_df) + geom_boxplot(aes(x = reorder(player, samples, median), samples, 
                               fill = reorder(player, samples, median)), show.legend=FALSE)
```
```{r}
theta_hat <- apply(fit2$theta, 2, mean)
ggplot(data.frame(size = tapply(white_df$White.ACPL, white_df$White.Player_ID, length), theta_hat = theta_hat), aes(size, theta_hat)) + geom_point()
```





