##Question 2 a
```{r}
chess_df <- read.csv("analysis.csv")
```
Pre-process data
```{r}
normalize<-function(data){
  return((data-mean(data)) / sd(data))
}

df <- chess_df[,c('Year','White.Num.Moves','Black.Num.Moves','Combined.ACPL','PreDeepBlue')]
df
process_df <- aggregate(x=df[c('Combined.ACPL','White.Num.Moves','Black.Num.Moves')],by=list(Year=df$Year,PreDeepBlue=df$PreDeepBlue),mean)
process_df
```
Explore the explanatory variables
```{r}
hist(process_df$Combined.ACPL)
```
Correlation of variables
```{r}
cor(process_df)
```
View relationship between Year and average combined ACPL
```{r}
plot(Combined.ACPL ~ Year, data = process_df)
```
View relationship between PreDeepBlue and average combined ACPL
```{r}
boxplot(Combined.ACPL ~ PreDeepBlue, data = process_df)
```
View relationship between White.Num.Moves and Black.Num.Moves and average combined ACPL
```{r}
pairs(subset(process_df, select = c(White.Num.Moves,Combined.ACPL,Black.Num.Moves)))
```
```{r}
process_df$PreDeepBlue <- factor(process_df$PreDeepBlue)
process_df
```
Model with 'PreDeepBlue' and 'Year'
```{r}
lm1 <- lm(Combined.ACPL ~ PreDeepBlue+Year, data = process_df)
summary(lm1)
```

```{r}
library(car)
qqPlot(lm1, which = 2)
```
```{r}
step_AIC_backward <- step(lm1)
step_AIC_backward
```
Re-fit model and consider the second order term
```{r}
lm2 <- lm(Combined.ACPL ~ Year+I(Year^2), data = process_df)
summary(lm2)
```

```{r}
qqPlot(lm2, which = 2)
```
```{r}
step_AIC_backward <- step(lm2)
step_AIC_backward
```
Predict players performance in next 10 years
```{r}
Year <- c(2022,2023,2024,2025,2026,2027,2028,2029,2030,2031)
pre_df <-  data.frame(Year)
pre_df
ACPL_predict <- predict(lm2, newdata = pre_df)
plot(pre_df$Year,ACPL_predict)
```

```{r}
ACPL_hat <- predict(lm2)
plot(process_df$Year,ACPL_hat)
```








